How to obtain the redshift distribution from probabilistic redshift estimates

### 2019-08-16 research/paper/draft.tex
filling in holes in intro

\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
\newcommand{\dwh}[1]{\todo{Attn. Hogg:}{blue}{#1}}
\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\mathul}[1]{\underline{#1}}
\newcommand{\Sect}[1]{Section~\ref{#1}}
\newcommand{\Eq}[1]{Equation~\ref{#1}}
\newcommand{\Fig}[1]{Figure~\ref{#1}}
\newcommand{\Tab}[1]{Table~\ref{#1}}
\newcommand{\project}[1]{{\textsc{#1}}}
\newcommand{\lsst}{\project{LSST}}
\newcommand{\desc}{\lsst-\project{DESC}}
\newcommand{\sdss}{\project{SDSS}}
\newcommand{\boss}{\project{BOSS}}
\newcommand{\des}{\project{DES}}
\newcommand{\Chippr}{\project{CHIPPR}}
\newcommand{\repo}[1]{{\texttt{#1}}~}
\newcommand{\qp}{\repo{qp}}
\newcommand{\chippr}{\repo{chippr}}
\newcommand{\cosmolike}{\repo{CosmoLike}}
\newcommand{\emcee}{\repo{emcee}}
\newcommand{\github}{\href{https://github.com}{GitHub}}
\newcommand{\python}{\textit{Python}}
\newcommand{\data}{\ensuremath{\vec{d}}}
% could change to bold
\newcommand{\like}{\mathscr{L}}
\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}
% could change to Prob or Pr 
\newcommand{\expect}[1]{\left<#1\right>}
\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
\newcommand{\gvn}{\mid}% could use | or \vert
\newcommand{\integral}[2]{\ensuremath{\int #1 \mathrm{d} #2}}
\newcommand{\sz}{spec-$z$}
\newcommand{\Sz}{Spec-$z$}
\newcommand{\pz}{photo-$z$}
\newcommand{\Pz}{Photo-$z$}
\newcommand{\pzpdf}{\pz\ PDF}
% could change to posterior
\newcommand{\Pzpdf}{\Pz\ PDF}
% could change to posterior
\newcommand{\zpdf}{redshift posterior}
% could change to implicit posterior
\newcommand{\pzip}{\pz implicit posterior}
\newcommand{\nz}{$n(z)$}
\newcommand{\Nz}{$N(z)$}
\newcommand{\stack}{$\hat{n}(z)$}
\newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
% could change to \vec
\newcommand{\ndphi}{\bvec{\phi}}
\newcommand{\mmle}{marginalized maximum likelihood estimate}
% really marginalized maximum a posteriori estimate, may still change this

@@

\author{Alex I. Malz\altaffilmark{1}}
\author{David W. Hogg\altaffilmark{1,2,3,4}}
\email{aimalz@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

@@

Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are not insignificant contributors to the total uncertainties reported on cosmological parameters.
As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
Based on the goals of a photometric galaxy survey, limits can be placed on the tolerance to these effects.
For example, the Science Requirements Document \citep{mandelbaum_weak_2017} states \lsst's requirements for the main cosmological sample, reproduced in Table~\ref{tab:lsstsrd}.

@@

\Pzpdf s have been produced by completed surveys \citep{hildebrandt_cfhtlens:_2012, sheldon_photometric_2012} and will be produced by ongoing and upcoming surveys \citep{abell_lsst_2009, carrasco_kind_exhausting_2014, bonnett_redshift_2016, masters_mapping_2015}.  
\Pzpdf s are not without their own shortcomings, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014, malz_approximating_2018} and the divergent results of each method used to derive them \citep[ and reviewed in Schmidt, Malz \& Soo, et al. (in prep)]{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.  
The most concerning weakness of \pzpdf s, however, is their usage in the literature, which is at best inconsistent and at worst incorrect.  
The most prevalent application of \pzpdf s is their use in estimating the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}, a quantity essential to the calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{mandelbaum_precision_2008, hildebrandt_cfhtlens:_2012, sheldon_photometric_2012, benjamin_cfhtlens_2013, kelly_weighing_2014, bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, leistedt_hierarchical_2016}, as well as to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.

@@

If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then the redshift PDFs are delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{j})$ in place of $\delta(z, z_{j}^{\dagger})$.
Though it is most common for $\hat{z}_{j}$ to be the maximum or \textit{mode} of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.

However, any procedure that reduces \pzpdf s to point estimates loses information about the uncertainty on redshift.
To circumvent this loss, it has grown to be more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
\begin{align}
\label{eqn:stack}
\hat{n}(z) &= \frac{1}{J} \sum_{j = 0}^{J} \pr{z}_{j}
\end{align}
for a sample of $J$ galaxies $j$, or, equivalently, the redshift distribution function $\hat{N}(z) = J \hat{n}(z)$, by effectively averaging the \pzpdf s.

%Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
%They have been used to form selection criteria of samples from galaxy surveys without propagation 
%through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
%Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.

Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
This paper challenges the logically invalid yet pervasive analysis procedure of stacking \pzpdf s by presenting and validating a hierarchical Bayesian technique for the use of \pzpdf s\ in the inference of \nz, yielding a method applicable to arbitrary one-point statistics relevant to cosmology, large-scale structure, and galaxy evolution; future work will extend this methodology to higher-order statistics.
We aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
Though others have approached the problem before \citep{leistedt_hierarchical_2016}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys for which there may be inertia preventing a complete restructuring of the analysis pipeline.

@@

\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzpdf\ codes.
Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
Template-fitting codes have an explicit prior input formed by redshifting a small number of templates, leading to a highly nonuniform but physically-motivated interim prior.
%Another potential method for selecting an interim prior with support over the entire redshift range expected of the photometric survey is to sum two or more $N(z)$ distributions obtained from reliable photometric surveys in the past.  
%This is just as problematic as using a biased spectroscopically derived $N(z)$ as the interim prior because the sum of redshift distributions for two or more surveys does not reflect our beliefs about the true distribution for a single survey even though it provides support over the same redshift range.  
%To simulate this case, we choose an interim prior with more weight at high and low redshifts than for mid-range redshifts.  
Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
% \aim{referenceh chapter 3 for complication of algorithm}
Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.  
%Because low-redshift galaxies are more likely to be bright enough to be observed by such a survey, $N(z)$ determined from that sample may be heavily biased to low redshift galaxies.  
%By contrast, the galaxies that were unobserved in such a survey are more likely be dimmer, making them more likely to be at higher redshifts.  
%Since the interim prior is not compatible with our beliefs about the true redshift distribution, the resulting interim redshift posteriors will be inappropriate.

@@

%In the tests on simulated data performed here, the full posterior distribution over the hyperparameters defining $N(z)$ derived by this method is consistent with the true redshift distribution function, making the mean of sampled values an excellent point estimator of $N(z)$.  
%The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.  
%\aim{Refer to quantitative results on KLD of \Nz\ and cosmological parameter space.}
%The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.  
%Tests were also performed on subsets of BOSS DR10 data with results consistent with those of simulations.

@@

%\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
%	if testable how so? outside the scope of paper, data producers be warned!
%	a likelihood is better -- give us that if you can!
%	focus on methods that acknowledge probabilistic structure of problem}

@@

\bibliographystyle{apj}
\bibliography{thesis}

>>>>

\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
\newcommand{\dwh}[1]{\todo{Attn. Hogg:}{blue}{#1}}

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\mathul}[1]{\underline{#1}}

\newcommand{\Sect}[1]{Section~\ref{#1}}
\newcommand{\Eq}[1]{Equation~\ref{#1}}
\newcommand{\Fig}[1]{Figure~\ref{#1}}
\newcommand{\Tab}[1]{Table~\ref{#1}}

\newcommand{\project}[1]{{\textsc{#1}}}
\newcommand{\lsst}{\project{LSST}}
\newcommand{\desc}{\lsst-\project{DESC}}
\newcommand{\sdss}{\project{SDSS}}
\newcommand{\boss}{\project{BOSS}}
\newcommand{\des}{\project{DES}}
\newcommand{\Chippr}{\project{CHIPPR}}% maybe change this to algorithm/italics

\newcommand{\repo}[1]{{\texttt{#1}}}
\newcommand{\qp}{\repo{qp}}
\newcommand{\chippr}{\repo{chippr}}
\newcommand{\cosmolike}{\repo{CosmoLike}}
\newcommand{\emcee}{\repo{emcee}}
\newcommand{\github}{\href{https://github.com}{GitHub}}% maybe remove link and use code font
\newcommand{\python}{\textit{Python}}% establish code font?

\newcommand{\data}{\ensuremath{\vec{d}}}% could change to bold
\newcommand{\like}{\mathscr{L}}
\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}% could change to Prob or Pr 
\newcommand{\expect}[1]{\left<#1\right>}
\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
\newcommand{\gvn}{\mid}% could use | or \vert
\newcommand{\integral}[2]{\ensuremath{\int #1 \mathrm{d} #2}}
\newcommand{\sz}{spec-$z$}
\newcommand{\Sz}{Spec-$z$}
\newcommand{\pz}{photo-$z$}
\newcommand{\Pz}{Photo-$z$}
\newcommand{\pzpdf}{\pz\ PDF}% could change to posterior
\newcommand{\Pzpdf}{\Pz\ PDF}% could change to posterior
\newcommand{\zpdf}{redshift posterior}% could change to implicit posterior
\newcommand{\pzip}{\pz implicit posterior}
\newcommand{\nz}{$n(z)$}
\newcommand{\Nz}{$N(z)$}
\newcommand{\stack}{$\hat{n}(z)$}
\newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}% could change to \vec
\newcommand{\ndphi}{\bvec{\phi}}
\newcommand{\mmle}{marginalized maximum likelihood estimate}% really marginalized maximum a posteriori estimate, may still change this

@@

\author{Alex I. Malz\altaffilmark{1,2}}
\author{David W. Hogg\altaffilmark{2,3,4,5}}
\email{aimalz@nyu.edu}

\altaffiltext{1}{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
\altaffiltext{3}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{4}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

@@

Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are a dominant contributor to the total uncertainties reported on cosmological parameters \citep{abruzzo_impact_2019}.
As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
Based on the goals of a photometric galaxy survey, limits can be placed on the tolerance to these effects.
For example, the Science Requirements Document \citep{mandelbaum_weak_2017} states \lsst's requirements for the main cosmological sample, reproduced in Table~\ref{tab:lsstsrd}.

@@

\Pzpdf s have been produced by completed surveys \citep{hildebrandt_cfhtlens:_2012, sheldon_photometric_2012} and will be produced by ongoing and upcoming surveys \citep{abell_lsst_2009, carrasco_kind_exhausting_2014, bonnett_redshift_2016, masters_mapping_2015}.  
\Pzpdf s are not without their own shortcomings, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014, malz_approximating_2018} and the divergent results of each method used to derive them \citep[ and reviewed in Schmidt, Malz \& Soo, et al. (in prep)]{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.  
The most concerning weakness of \pzpdf s, however, is their usage in the literature, which is at best inconsistent and at worst incorrect.

If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then the redshift PDFs are delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{j})$ in place of $\delta(z, z_{j}^{\dagger})$.
Though it is most common for $\hat{z}_{j}$ to be the maximum or \textit{mode} of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.
However, any procedure that reduces \pzpdf s to point estimates discards valuable information about the uncertainty on redshift.

Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
They have been used to form selection criteria of samples from galaxy surveys without propagation through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.

The most prevalent application of \pzpdf s that preserves their information content is the estimation of the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}.
The redshift distribution is a key input to the traditional calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{mandelbaum_precision_2008, hildebrandt_cfhtlens:_2012, sheldon_photometric_2012, benjamin_cfhtlens_2013, kelly_weighing_2014, bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, yang_calibrating_2018}, as well as to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.
Even with \pz s adhering to the \lsst\ requirements of \Tab{tab:lsstsrd}, the accuracy and precision to which \nz\ must be known limit the degree to which constraints on the cosmological parameters can advance \citep{abruzzo_impact_2019}.

@@

Though it is traditional to estimate \nz\ from \pz\ point estimates \citep{abruzzo_impact_2019}, there is a great deal of literature employing \pzpdf s.
To circumvent this loss, it has become to be more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
\begin{align}
\label{eqn:stack}
\hat{n}(z) &= \frac{1}{J} \sum_{j = 0}^{J} \pr{z}_{j}
\end{align}
for a sample of $J$ galaxies $j$, or, equivalently, the redshift distribution function $\hat{N}(z) = J \hat{n}(z)$, by effectively averaging the \pzpdf s.

Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
This paper challenges the logically invalid yet pervasive analysis procedure of stacking \pzpdf s by presenting and validating a hierarchical Bayesian technique for the use of \pzpdf s\ in the inference of \nz, yielding a method applicable to arbitrary one-point statistics relevant to cosmology, large-scale structure, and galaxy evolution; future work will extend this methodology to higher-order statistics.
We aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
Though others have approached the problem before \citep{leistedt_hierarchical_2016, leistedt_hierarchical_2018}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys for which there may be inertia preventing a complete restructuring of the analysis pipeline.

@@

\chippr\ can handle any implicit prior with support over the redshift range where \nz\ is defined, but some archetypes of implicit prior are more likely to be encountered in the wilds of \pzpdf\ codes.
Ideally, an uninformative implicit prior would be used, although it may be complicated to compute from the covariances of the raw data.  
Template-fitting codes have an explicit prior input formed by redshifting a small number of templates, leading to a highly nonuniform but physically-motivated interim prior.
%Another potential method for selecting an interim prior with support over the entire redshift range expected of the photometric survey is to sum two or more $N(z)$ distributions obtained from reliable photometric surveys in the past.  
%This is just as problematic as using a biased spectroscopically derived $N(z)$ as the interim prior because the sum of redshift distributions for two or more surveys does not reflect our beliefs about the true distribution for a single survey even though it provides support over the same redshift range.  
%To simulate this case, we choose an interim prior with more weight at high and low redshifts than for mid-range redshifts.  
Machine learning approaches tend to be trained on previously observed data sets that are biased towards low redshift, which biases the implicit prior towards low redshift.
% \aim{reference chapter 3 for complication of algorithm}
Some efforts have been made to modify an observationally informed implicit prior so that it is more representative of the photometric data for which redshifts are desired \citep{sheldon_photometric_2012}, but, unless it is equal to the true \nz, it will propagate to the results of traditional \nz\ estimation methods.  
%Because low-redshift galaxies are more likely to be bright enough to be observed by such a survey, $N(z)$ determined from that sample may be heavily biased to low redshift galaxies.  
%By contrast, the galaxies that were unobserved in such a survey are more likely be dimmer, making them more likely to be at higher redshifts.  
%Since the interim prior is not compatible with our beliefs about the true redshift distribution, the resulting interim redshift posteriors will be inappropriate.

@@

%In the tests on simulated data performed here, the full posterior distribution over the hyperparameters defining $N(z)$ derived by this method is consistent with the true redshift distribution function, making the mean of sampled values an excellent point estimator of $N(z)$.  
%The information contained in the full posterior distribution's shape convey the traditional error bar information without having to explicitly propagate any error estimates.  
\aim{TODO: Refer to quantitative results on KLD of \Nz\ and cosmological parameter space.}
%The results of those tests is summarized below and in \Tab{tab:kld}, where lower values indicate a closer match between the true $N(z)$ and the estimator.  
%Tests were also performed on subsets of BOSS DR10 data with results consistent with those of simulations.

@@

\aim{TODO: new paragraph for what can go wrong, call to community for what to do about it, what aspects of implicit prior are knowable and not knowable?
	if testable how so? outside the scope of paper, data producers be warned!
	a likelihood is better -- give us that if you can!
	focus on methods that acknowledge probabilistic structure of problem}

@@

\bibliographystyle{apj}
\bibliography{draft}
